```{r}
library(tidyverse)
library(SingleCellExperiment)
library(Matrix)
library(Matrix.utils)
library(corrplot)
library(gplots)
library(edgeR)
library(dendextend)
library(vroom)
library(patchwork)
```

## Hierarchical clustering of annotated cell types
```{r fig.height=6, fig.width=6}
eb_pseudobulk_counts <- vroom("data/static_qtl_calling/eb_cellid/pseudobulk_tmm/basic/pseudobulk_all.tsv") %>%
  column_to_rownames("gene") %>% t %>% as("dgCMatrix")

eb_pseudobulk_counts_agg <- aggregate.Matrix(eb_pseudobulk_counts, 
                                             groupings=str_extract(rownames(eb_pseudobulk_counts), "[^_]+$"), 
                                             fun="sum")

gtf_pcgs <- vroom("data/gencode/gencode.hg38.filtered.gtf")
pcgs <- gtf_pcgs$hgnc
hvgs <- read_tsv("data/fca/fca_subsampled_hvg.tsv")

keeper_genes <- intersect(intersect(colnames(eb_pseudobulk_counts_agg), pcgs), hvgs$gene)
eb_pseudobulk_counts_agg <- eb_pseudobulk_counts_agg[,keeper_genes]
eb_pseudobulk_norm <- DGEList(counts=t(eb_pseudobulk_counts_agg)) %>% 
  calcNormFactors(method="TMM") %>% 
  edgeR::cpm(log=T)

hierarchical_clustering <- hclust(dist(t(eb_pseudobulk_norm)), method="ward.D2")
plot(hierarchical_clustering)

type_order <- hierarchical_clustering$labels[hierarchical_clustering$order]
```

```{r fig.height=8, fig.width=10}
avg_dend_obj <- as.dendrogram(hierarchical_clustering)
avg_col_dend <- color_branches(avg_dend_obj, col=c("#004D80", "#00AB8E", "#970E53", "#0076BA", "#FEAE00", "#965C3D", "#B51700"), h = 125)

rotated_order <- c("Neuroendocrine-cells", "PNS-neurons", "Skeletal-muscle-cells", 
                   "CNS-neurons" , "Retinal-cells", "CNS-glia",
                   "Metanephric-cells", "Ciliated-epithelial-cells",
                   "Acinar-cells", "Ductal-cells", "Erythroblasts",
                   "PNS-glia",  "Lymphoid-cells", "Megakaryocytes",
                   "Bronchiolar-and-alveolar-epithelial-cells", "Adrenocortical-cells",
                   "Parietal-and-chief-cells", "Ureteric-bud-cells",
                   "Vascular-endothelial-cells",
                   "Epicardial-fat-cells", "Mesothelial-cells", "Cardiomyocytes", "Mesangial-cells",
                   "Stellate-cells", "Smooth-muscle-cells", "Stromal-cells",
                   "Squamous-epithelial-cells", "Goblet-cells", "Hepatoblasts")  
```

```{r fig.height=4, fig.width=5}
par(mar = c(2,2,2,40))
rotate(avg_col_dend, order=rotated_order) %>% 
  set("labels_cex", 1.5) %>%
  plot(horiz=T)
```

## UMAP Embedding
Load the UMAP embedding and cell type assignments
```{r}
umap <- vroom("data/single_cell_objects/eb_pflog1ppfnorm.hvg.umap_embedding.txt", col_names=c("UMAP_1", "UMAP_2"))
celltypes <- vroom("data/fca/eb_cellid_labels.tsv")

sc_annotated <- celltypes %>%
  mutate(umap1=umap$UMAP_1, umap2=umap$UMAP_2)
```

Assign colors for each cell type in following with the hierarchical clustering above
```{r}
type_order_with_unassigned <- c(rotated_order, "unassigned")
type_colors <- c("#130063", "#192594", "#004D80", 
                 "#7F7FFF", "#316DFF", "#1B7FFF", "#0097FF", "#7298BB", 
                 "#7F9ACE", "#8CB4D5", "#96C0E5", "#98BEC6", 
                 "#00AB8E", "#3EBFBB", "#55D0B7", "#25C683", 
                 "#860B1E", "#9B1716", "#B51700", "#B13B3E", 
                 "#CC4E4F", "#D32323", "#FF0000",
                 "#965C3D", "#8410D3", "#84108A",  
                 "#FEDE7F", "#FED255", "#FEAE00", 
                 "#D5D5D5") 

color_map <- tibble(type = type_order_with_unassigned, color = type_colors) %>%
  mutate(type_spaced = str_replace_all(type, "-", " "))
```

```{r fig.width=8, fig.height=4}
umap_viz <- sc_annotated %>%
  mutate(value = factor(value, levels=color_map$type_spaced)) %>%
  drop_na() # removes a few very rare cell types in the EB data

umap_plot <- ggplot(umap_viz, aes(x=umap1, y=umap2, color=value)) +
  geom_point(size=1e-5) +
  theme_classic(base_size=20) +
  scale_color_manual(values = color_map$color) + 
  theme(axis.ticks.x = element_blank(), axis.text.x=element_blank(),
        axis.ticks.y=element_blank(), axis.text.y=element_blank(),
        legend.position="none") +
  xlab("UMAP1") + ylab("UMAP2")
```

### Cell reassignment
How many of the unassigned cells are re-annotated using a classifier that includes an IPSC label?
```{r}
celltypes_ipsc <- vroom("data/fca/eb_cellid_labels.with_ipsc.tsv")

reassigned_cells <- filter(celltypes, value=="unassigned") %>%
  left_join(celltypes_ipsc, by="cell")

sum(reassigned_cells$value.y=="IPSC") / nrow(reassigned_cells)
```


## Overview of Pseudobulk Data
We see that the proportion of cells mapping to each cell type varies widely between donors
```{r, fig.height=8, fig.width=8}
cell_metadata <- vroom("data/single_cell_objects/highpass/eb_metadata.tsv")

donor_celltype <- dplyr::select(cell_metadata, c("cell", "donor_id")) %>%
  inner_join(dplyr::select(sc_annotated, c("cell", "value"))) 
```

Here's what it looks like for different cutoffs on the number of cells req'd for a 
pseudobulk sample
```{r}
thresholds <- seq(10) * 5
all_donors <- unique(donor_celltype$donor_id)

count_celltypes <- function(df, thresh, d) {
  dplyr::count(df, donor_id, value) %>% filter((n>=thresh) & (donor_id==d)) %>% nrow
}

n_types <- expand_grid(thresholds, all_donors) %>%
  mutate(n_types=map2_dbl(thresholds, all_donors, count_celltypes, df=donor_celltype))

ggplot(n_types, aes(x=thresholds, y=n_types, color=all_donors)) +
  geom_line()
```

But most donors to generate most cell types
```{r fig.height=4, fig.width=3}
atleast_5cells <- filter(n_types, thresholds==5)  %>%
  mutate(all_donors = factor(all_donors, levels=pull(arrange(., n_types), all_donors)))
donor_order <- levels(atleast_5cells$all_donors)

types_plot <- ggplot(atleast_5cells, aes(x=all_donors, y=n_types)) +
  geom_segment( aes(x=all_donors, xend=all_donors, y=0, yend=n_types), color="skyblue") +
  geom_point( color="blue", size=5, alpha=0.8) +
  geom_hline(linetype="dashed", color="red", yintercept=median(atleast_5cells$n_types)) +
  theme_classic(base_size=18) +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  xlab("Donor") + ylab("Number of Cell Types")
```

```{r}
proportions_per_celltype <- donor_celltype %>%
  add_count(donor_id, name="donor_n") %>%
  add_count(donor_id, value, name="donor_type_n") %>%
  mutate(proportion=donor_type_n / donor_n) %>%
  dplyr::select(value, donor_id, proportion) %>%
  distinct %>%
  mutate(value = factor(value, levels=color_map$type_spaced)) %>%
  drop_na() %>% # removes rare cell types
  mutate(donor_id=factor(donor_id, levels=donor_order))

props_plot <- ggplot(proportions_per_celltype, aes(x=donor_id, y=proportion, fill=value)) +
  geom_bar(stat="identity", color="black") +
  theme_classic(base_size=18) +
  theme(axis.text.x=element_text(angle=45, hjust=1)) +
  scale_fill_manual(values=color_map$color) +
  xlab("Donor") + ylab("Cell Type Proportion") + 
  theme(legend.position="none")
props_plot
```

Stitch them all together
```{r fig.width=14, fig.height=8}
layout <- "
##AAA###
##AAA###
##AAA###
##AAA###
##AAA###
##AAA###
BBBBCCCC
BBBBCCCC
BBBBCCCC
"
umap_plot + types_plot + props_plot +
  plot_layout(design = layout)

```

