```{r}
library(tidyverse)
```

## Overview of single-cell data
Load the UMAP embedding and cell type assignments
```{r}
umap <- vroom("data/single_cell_objects/eb_pflog1ppfnorm.hvg.umap_embedding.txt", col_names=c("UMAP_1", "UMAP_2"))
celltypes <- vroom("data/fca/eb_cellid_labels.tsv")

sc_annotated <- celltypes %>%
  mutate(umap1=umap$UMAP_1, umap2=umap$UMAP_2)
```

Stratify by germ layer
```{r}
dropped <- c("STC2_TLX1 positive cells","Lymphatic endothelial cells",
             "Intestinal epithelial cells","Endocardial cells")
ecto <- c("Retinal cells","CNS neurons","PNS glia","CNS glia","PNS neurons",
          "Adrenocortical cells","Neuroendocrine cells")
immune <- c("Erythroblasts","Megakaryocytes","Lymphoid cells")
meso <- c("Vascular endothelial cells","Stromal cells","Epicardial fat cells",
          "Skeletal muscle cells","Mesangial cells","Mesothelial cells",
          "Cardiomyocytes","Smooth muscle cells","Stellate cells")
epi <- c("Ductal cells", "Metanephric cells","Ciliated epithelial cells",
         "Ureteric bud cells","Acinar cells",
         "Parietal and chief cells")
endo <- c("Hepatoblasts","Goblet cells","Bronchiolar and alveolar epithelial cells",
          "Squamous epithelial cells")

sc_annotated <- sc_annotated %>%
  mutate(ecto_type=if_else(value %in% ecto, value, NA_character_)) %>%
  mutate(endo_type=if_else(value %in% endo, value, NA_character_)) %>%
  mutate(meso_type=if_else(value %in% meso, value, NA_character_)) %>%
  mutate(epi_type=if_else(value %in% epi, value, NA_character_)) %>%
  mutate(immune_type=if_else(value %in% immune, value, NA_character_))
```

Ectoderm cells
```{r}
ggplot(sc_annotated, aes(x=umap1, y=umap2, color=ecto_type)) +
  geom_point(pch=".") +
  theme_classic()
```

Mesoderm cells
```{r}
ggplot(sc_annotated, aes(x=umap1, y=umap2, color=meso_type)) +
  geom_point(pch=".") +
  theme_classic(base_size=12) +
  guides(color=guide_legend(override.aes=list(size=10)))
```

Endoderm cells
```{r}
ggplot(sc_annotated, aes(x=umap1, y=umap2, color=endo_type)) +
  geom_point(pch=".") +
  theme_classic()
```

Epithelial cells
```{r}
ggplot(sc_annotated, aes(x=umap1, y=umap2, color=epi_type)) +
  geom_point(pch=".") +
  theme_classic()
```

Immune cells
```{r}
ggplot(sc_annotated, aes(x=umap1, y=umap2, color=immune_type)) +
  geom_point(pch=".") +
  theme_classic()
```

## Overview of Pseudobulk Data
First, which cell types are included in this analysis (after filtering out samples with too few cells)?
```{r, warning=FALSE, message=FALSE}
options(readr.show_progress = FALSE)
sample_summary <- read_tsv("/project2/gilad/jpopp/ebQTL/data/static_qtl_calling/eb_cellid/pseudobulk_tmm/sample_summary.tsv") %>%
  mutate(individual=factor(individual), type=factor(type))

cell.types <- sample_summary  %>%
  mutate(type=as.character(type)) %>%
  filter(!dropped) %>%
  dplyr::count(type) %>%
  filter(n > 25) %>%
  pull(type)

cell.types
```

How many samples do we have per cell type?
```{r}
sample_summary  %>%
  filter(!dropped) %>%
  select(individual, type, dropped) %>%
  pivot_wider(names_from=type, values_from=dropped)
  ggplot(aes(x=type, y=n, fill=type)) +
  geom_bar(stat="identity") +
  ylab("n samples") +
  theme_classic(base_size=15) +
  theme(axis.text.x = element_text(angle = 45, hjust=1), legend.position="none") +
  geom_abline(slope=0, intercept=25, linetype="dashed")
```

To emphasize that we do see most cell types in most donors, we can visualize this differently
```{r, fig.width=8, fig.height=4}
read_tsv("/project2/gilad/jpopp/ebQTL/data/static_qtl_calling/eb_cellid/pseudobulk_tmm/sample_summary.tsv") %>%
  mutate(individual=factor(individual), type=factor(type))  %>%
  #filter(!dropped) %>%
  select(individual, type, dropped) %>%
  pivot_wider(names_from=type, values_from=dropped, values_fill=T) %>%
  pivot_longer(!individual, names_to="type", values_to="dropped") %>%
  mutate(individual=paste0("NA", individual)) %>%
  mutate(observed=!dropped) %>%
  ggplot(aes(x = individual, y = type, color = factor(observed, levels=c(T, F)))) +
    geom_point(shape = 1, size = 1, stroke=2) +
    scale_color_manual(values = c("#1DB100", "#D5D5D5")) +
    theme_classic(base_size=15) +
    labs(title = "Cell Type Observations",
         x = "Donor",
         y = "Cell Type",
         color = "Cell Type Observed") +
    theme(axis.text.x = element_text(angle = 45, hjust=1))
```

After manual pruning of outlier samples
```{r, fig.width=8, fig.height=4}
read_tsv("/project2/gilad/jpopp/ebQTL/data/static_qtl_calling/eb_cellid/pseudobulk_tmm/basic/sample_summary_manual.tsv") %>%
  mutate(individual=factor(individual), type=factor(type))  %>%
  #filter(!dropped) %>%
  select(individual, type, dropped) %>%
  pivot_wider(names_from=type, values_from=dropped, values_fill=T) %>%
  pivot_longer(!individual, names_to="type", values_to="dropped") %>%
  mutate(individual=paste0("NA", individual)) %>%
  mutate(observed=!dropped) %>%
  ggplot(aes(x = individual, y = type, color = factor(observed, levels=c(T, F)))) +
    geom_point(shape = 1, size = 1, stroke=2) +
    scale_color_manual(values = c("#1DB100", "#D5D5D5")) +
    theme_classic(base_size=15) +
    labs(title = "Cell Type Observations",
         x = "Donor",
         y = "Cell Type",
         color = "Cell Type Observed") +
    theme(axis.text.x = element_text(angle = 45, hjust=1))
```

How many cells do we have (median) per cell type?
```{r}
sample_summary %>%
  filter(!dropped) %>%
  filter(type %in% cell.types) %>%
  group_by(type) %>%
  summarize(n_cells_median=median(n_cells_filtered)) %>%
  arrange(n_cells_median) %>%
  ggplot(aes(x=type, y=n_cells_median)) +
  geom_bar(stat="identity") +
  theme(axis.text.x = element_text(angle = 45, hjust=1))
```

And how many cell types per individual?
```{r}
sample_summary  %>%
  filter(!dropped) %>%
  dplyr::count(individual) %>%
  ggplot(aes(x=individual, y=n, fill=individual)) +
  geom_bar(stat="identity") +
  ylab("n samples") +
  theme_classic(base_size=15) +
  theme(axis.text.x = element_text(angle = 45, hjust=1), legend.position="none") +
  geom_abline(slope=0, intercept=25, linetype="dashed")
```
