---
title: "MASH"
output: html_notebook
---
```{r, echo=FALSE, results=FALSE, warning=FALSE}
library(tidyverse)
library(gplots)
library(mashr)
library(rmeta)
library(vroom)
library(patchwork)
```

# Characterizing mash patterns
```{r}
load("/project2/gilad/jpopp/ebQTL/results/static_qtl_calling/eb_cellid/pseudobulk_tmm/basic/8pcs/mash_inputs.Rdata")
```

For these visualizations, I'm going to keep the cell type order established
by hierarchical clustering in the `analysis/annotation/cell_type_annotation.Rmd` notebook
```{r}
color_map <- readRDS("/project2/gilad/jpopp/ebQTL/temp/color_map.rds")
```


Component 1 - dense factor
```{r}
factor_1 <- as_tibble(princomp(U.flash$FLASH_nonneg_1)$loadings[,1], rownames="type") %>%
  left_join(color_map, by="type") %>%
  arrange(type) %>%
  mutate(type=factor(type, levels=color_map$type))
```

Component 2 - mesoderm/ endoderm
```{r}
factor_2 <- as_tibble(princomp(U.flash$FLASH_nonneg_2)$loadings[,1], rownames="type") %>%
  left_join(color_map, by="type") %>%
  arrange(type) %>%
  mutate(type=factor(type, levels=color_map$type))
```

FLASH 3 - Most differentiated cell types
```{r}
factor_3 <- as_tibble(princomp(U.flash$FLASH_nonneg_3)$loadings[,1], rownames="type") %>%
  left_join(color_map, by="type") %>%
  arrange(type) %>%
  mutate(type=factor(type, levels=color_map$type))
```

FLASH 4 - Neuroepithelium
```{r}
factor_4 <- as_tibble(princomp(U.flash$FLASH_nonneg_4)$loadings[,1], rownames="type") %>%
  left_join(color_map, by="type") %>%
  arrange(type) %>%
  mutate(type=factor(type, levels=color_map$type))
```

Visualize patterns
```{r fig.width=8, fig.height=4}
factors_all <- bind_rows(factor_1, factor_2, factor_3, factor_4) %>%
  mutate(factor_name=rep(c("Factor 1", "Factor 2", "Factor 3", "Factor 4"), each=29)) 

mash_factors_all <- ggplot(factors_all, aes(x=type_spaced, y=value, fill=color)) +
    geom_bar(stat="identity") +
    theme_classic(base_size=20) +
    theme(axis.text.x = element_text(angle=30, hjust=1), 
          legend.position="none") +
    xlab("Cell Type") + ylab("Weight") +
    scale_fill_identity() +
    coord_flip() +
    lims(x = color_map$type_spaced) +
  facet_grid(cols=vars(factor_name))

mash_factors_all
```

## Exploring Specific Examples
```{r}
full_m <- readRDS("/project2/gilad/jpopp/ebQTL/results/static_qtl_calling/eb_cellid/pseudobulk_tmm/basic/8pcs/mash_fitted_model.full.rds")
```

```{r}
mash_colors <- p
```

```{r}
mash_plot_meta(full_m, "SH3PXD2B_rs10042482", labels=colnames(full_m$result$lfsr),
               colors=meta.colors(box=color_map$color))
```

How does this compare to the effects we learned in the celltype-by-celltype analysis?
```{r fig.width=4, fig.height=5}
nonmash_effects <- vroom("results/static_qtl_calling/eb_cellid/pseudobulk_tmm/basic/8pcs/tensorqtl_nominal.all.tsv")
nonmash_viz <- filter(nonmash_effects, (variant_id=="rs10042482") & (phenotype_id=="SH3PXD2B"))
```

Get the number of samples from this cell type to translate standard errors into standard deviation for consistency with mash
```{r}
sample_summary <- vroom("data/static_qtl_calling/eb_cellid/pseudobulk_tmm/basic/sample_summary_manual.tsv")
nsamples <- sample_summary %>% filter(!dropped) %>% group_by(type) %>% dplyr::count()
```


```{r}
mash_posterior_betas <- get_pm(full_m)["SH3PXD2B_rs10042482",]
mash_posterior_sds <- get_psd(full_m)["SH3PXD2B_rs10042482",]
mash_precision <- 1 / mash_posterior_sds ** 2
mash_celltypes <- colnames(get_pm(full_m))
mash_ests <- tibble(type=mash_celltypes,
                    beta=mash_posterior_betas,
                    sd=mash_posterior_sds,
                    precision=mash_precision,
                    approach=rep("mash", length(mash_celltypes)))

nonmash_betas <- nonmash_viz$slope
nonmash_ses <- nonmash_viz$slope_se 
nonmash_celltypes <- nonmash_viz$celltype
nonmash_ests <- tibble(type=nonmash_celltypes,
                    beta=nonmash_betas,
                    se=nonmash_ses,
                    approach=rep("non-mash", length(nonmash_celltypes))) %>%
  left_join(nsamples, by="type") %>%
  mutate(sd = nonmash_ses * sqrt(n)) %>%
  mutate(precision = 1 / nonmash_sds ** 2) %>%
  dplyr::select(type, beta, sd, precision, approach)
  
mash_comp_viz <- bind_rows(mash_ests, nonmash_ests) %>%
  mutate(approach=case_match(factor(approach, levels=c("non-mash", "mash")), "non-mash" ~ "Univariate Analysis", "mash" ~ "Multivariate Analysis")) %>%
  mutate(approach=factor(approach, levels=c("Univariate Analysis", "Multivariate Analysis"))) %>%
  left_join(color_map, by="type") %>%
  mutate(type_spaced=factor(type_spaced, levels=color_map$type_spaced))

univariate_only_viz <- nonmash_ests %>%
  left_join(color_map, by="type") %>%
  mutate(type_spaced=factor(type_spaced, levels=color_map$type_spaced))

univariate_only_example_plot <- ggplot(univariate_only_viz, aes(x=type_spaced, y=beta, color=color)) +
  geom_hline(aes(yintercept=0), linetype="dashed", color="red") +
  geom_errorbar(aes(ymin=beta - sd, ymax=beta + sd), color="#D5D5D5", width=0.5) +
  geom_point(aes(size=precision), shape=15) +
  theme_classic(base_size=20) +
  theme(legend.position = "none", axis.text.x = element_text(angle=45, hjust=1)) +
  scale_color_identity() +
  coord_flip() +
  xlab("Cell Type") +
  ylab("Estimated Effect Size")
```

```{r}
mash_only_viz <- mash_ests %>%
  left_join(color_map, by="type") %>%
  mutate(type_spaced=factor(type_spaced, levels=color_map$type_spaced))

mash_only_example_plot <- ggplot(mash_only_viz, aes(x=type_spaced, y=beta, color=color)) +
  geom_hline(aes(yintercept=0), linetype="dashed", color="red") +
  geom_errorbar(aes(ymin=beta - sd, ymax=beta + sd), color="#D5D5D5", width=0.5) +
  geom_point(aes(size=precision), shape=15) +
  theme_classic(base_size=20) +
  theme(legend.position = "none", axis.text.x = element_text(angle=45, hjust=1)) +
  scale_color_identity() +
  coord_flip() +
  xlab("Cell Type") +
  ylab("Estimated Effect Size")
```

Make a version of the mash plot without cell type labels (for graphical abstract)
```{r}
mash_only_example_plot <- ggplot(mash_only_viz, aes(x=type_spaced, y=beta, color=color)) +
  geom_hline(aes(yintercept=0), linetype="dashed", color="red") +
  geom_errorbar(aes(ymin=beta - sd, ymax=beta + sd), color="#D5D5D5", width=0.5) +
  geom_point(aes(size=precision), shape=15) +
  theme_classic(base_size=20) +
  theme(legend.position = "none", axis.text.x = element_text(angle=45, hjust=1), axis.text.y = element_blank(), axis.ticks.y = element_blank()) +
  scale_color_identity() +
  coord_flip() +
  xlab("Cell Type") +
  ylab("Estimated Effect Size")
```


```{r fig.height=4, fig.width=10}
png(filename="/project2/gilad/jpopp/ebQTL/figs/fig2/mash.png", width=1900, height=750)
layout <- "
  AAAABBB
  AAAABBB
"
mash_factors_all + mash_only_example_plot +
  plot_layout(design=layout)
dev.off()

png(filename="/project2/gilad/jpopp/ebQTL/figs/fig2/univariate.png", width=900, height=750)
univariate_only_example_plot
dev.off()
```

## Comparing weights
```{r, fig.width=6, fig.height=8}
fitted_m <- readRDS("/project2/gilad/jpopp/ebQTL/results/static_qtl_calling/eb_cellid/pseudobulk_tmm/basic/8pcs/mash_trained_model.rds")
```

```{r}
as_tibble(get_estimated_pi(full_m), rownames="component") %>%
  mutate(component=factor(component, levels=names(get_estimated_pi(full_m)))) %>%
  ggplot(aes(x=component, y=value)) +
  geom_bar(stat="identity") +
  scale_x_discrete(limits=rev) +
  xlab("Regulatory Pattern") +
  ylab("Weight") +
  coord_flip() +
  theme_classic(base_size = 10)
```

# Comparing mash QTLs to single cell type analysis
```{r}
mash_n_signif <- as_tibble(get_n_significant_conditions(full_m, thresh=0.05, sig_fn=get_lfsr), rownames="gv") %>%
  separate(gv, into=c("EB_HGNC", "EB_VARIANT_ID"), sep="_")

celltype_qtls_loc <- "/project2/gilad/jpopp/ebQTL/results/static_eqtl_followup/qtl_sets/tensorqtl/original/signif_variant_gene_pairs.bed"

hdc_distinct <- vroom(celltype_qtls_loc) %>% 
  dplyr::select(c(EB_HGNC, EB_VARIANT_ID)) %>%
  distinct()

mash_distinct <- mash_n_signif %>%
  filter(value > 0) %>%
  dplyr::select(c(EB_HGNC, EB_VARIANT_ID)) %>%
  distinct()

mash_novel <- anti_join(mash_distinct, hdc_distinct)

nrow(mash_novel)
length(unique(mash_novel$EB_HGNC))
```

