---
title: "MASH"
output: html_notebook
---
```{r, echo=FALSE, results=FALSE, warning=FALSE}
library(tidyverse)
library(gplots)
library(mashr)
library(rmeta)
library(vroom)
library(patchwork)
```

# Characterizing mash patterns
```{r}
load("/project2/gilad/jpopp/ebQTL/results/static_qtl_calling/eb_cellid/pseudobulk_tmm/basic/8pcs/mash_inputs.Rdata")
```

For these visualizations, I'm going to keep the cell type order established
by hierarchical clustering in the `analysis/annotation/tidy_celltype_names.Rmd` notebook
```{r}
type_levels <- c("Neuroendocrine-cells", "PNS-neurons", "Skeletal-muscle-cells",                               
                   "PNS-glia", "CNS-neurons" , "Retinal-cells", "CNS-glia",
                   "Metanephric-cells", "Erythroblasts", "Ciliated-epithelial-cells",
                   "Acinar-cells", "Ductal-cells", "Bronchiolar-and-alveolar-epithelial-cells", "Adrenocortical-cells",
                   "Parietal-and-chief-cells", "Ureteric-bud-cells",
                   "Epicardial-fat-cells", "Mesothelial-cells", "Cardiomyocytes", "Mesangial-cells",
                   "Stellate-cells", "Smooth-muscle-cells", "Stromal-cells",
                   "Vascular-endothelial-cells", "Lymphoid-cells", "Megakaryocytes",
                   "Squamous-epithelial-cells", "Goblet-cells", "Hepatoblasts")  

ectoderm_1 <- tibble(type=c("Neuroendocrine-cells", "PNS-neurons", "Skeletal-muscle-cells"),
                     layer="ecto_1", color="#004D80")
ectoderm_2 <- tibble(type=c("PNS-glia", "CNS-neurons" , "Retinal-cells", "CNS-glia"),
                     layer="ecto_2", color="#0076BA")
neuroepithelium <- tibble(type=c("Metanephric-cells", "Erythroblasts", "Ciliated-epithelial-cells",
                                 "Acinar-cells", "Ductal-cells", "Bronchiolar-and-alveolar-epithelial-cells", "Adrenocortical-cells",
                                 "Parietal-and-chief-cells", "Ureteric-bud-cells"),
                          layer="neuroepithelium", color="#00AB8E")
mesoderm <- tibble(type=c("Epicardial-fat-cells", "Mesothelial-cells", "Cardiomyocytes", "Mesangial-cells",
                          "Stellate-cells", "Smooth-muscle-cells", "Stromal-cells"),
                   layer="mesoderm", color="#B51700")
vasc <- tibble(type=c("Vascular-endothelial-cells"),
               layer="vasc", color="#965C3D")
immune <- tibble(type=c("Lymphoid-cells", "Megakaryocytes"),
                 layer="immune", color="#970E53")
endoderm <- tibble(type=c("Squamous-epithelial-cells", "Goblet-cells", "Hepatoblasts"),
                   layer="endoderm", color="#FEAE00")

color_map <- bind_rows(ectoderm_1, ectoderm_2, neuroepithelium, mesoderm, vasc, immune, endoderm)
```

Component 1 - dense factor
```{r}
factor_1 <- as_tibble(princomp(U.flash$FLASH_nonneg_1)$loadings[,1], rownames="type") %>%
  left_join(color_map, by="type") %>%
  arrange(type) %>%
  mutate(type=factor(type, levels=color_map$type))
```

Component 2 - mesoderm/ endoderm
```{r}
factor_2 <- as_tibble(princomp(U.flash$FLASH_nonneg_2)$loadings[,1], rownames="type") %>%
  left_join(color_map, by="type") %>%
  arrange(type) %>%
  mutate(type=factor(type, levels=color_map$type))
```

FLASH 3 - Most differentiated cell types
```{r}
factor_3 <- as_tibble(princomp(U.flash$FLASH_nonneg_3)$loadings[,1], rownames="type") %>%
  left_join(color_map, by="type") %>%
  arrange(type) %>%
  mutate(type=factor(type, levels=color_map$type))
```

FLASH 4 - Neuroepithelium
```{r}
factor_4 <- as_tibble(princomp(U.flash$FLASH_nonneg_4)$loadings[,1], rownames="type") %>%
  left_join(color_map, by="type") %>%
  arrange(type) %>%
  mutate(type=factor(type, levels=color_map$type))
```

Visualize patterns
```{r}
factors_all <- bind_rows(factor_1, factor_2, factor_3, factor_4) %>%
  mutate(factor_name=rep(c("Factor 1", "Factor 2", "Factor 3", "Factor 4"), each=29)) 

mash_factors_all <- ggplot(factors_all, aes(x=type, y=value, fill=color)) +
    geom_bar(stat="identity") +
    theme_classic(base_size=20) +
    theme(axis.text.x = element_text(angle=30, hjust=1), 
          legend.position="none") +
    xlab("Cell Type") + ylab("Weight") +
    scale_fill_identity() +
    coord_flip() +
    lims(x = rev(type_order)) +
  facet_grid(cols=vars(factor_name))

mash_factors_all
```

## Comparing weights
```{r, fig.width=6, fig.height=8}
fitted_m <- readRDS("/project2/gilad/jpopp/ebQTL/results/static_qtl_calling/eb_cellid/pseudobulk_tmm/basic/8pcs/mash_trained_model.rds")
```

```{r}
as_tibble(get_estimated_pi(fitted_m), rownames="component") %>%
  mutate(component=factor(component, levels=names(get_estimated_pi(fitted_m)))) %>%
  ggplot(aes(x=component, y=value)) +
  geom_bar(stat="identity") +
  scale_x_discrete(limits=rev) +
  xlab("Regulatory Pattern") +
  ylab("Weight") +
  coord_flip() +
  theme_classic(base_size = 10)
```



## Exploring Specific Examples
```{r}
full_m <- readRDS("/project2/gilad/jpopp/ebQTL/results/static_qtl_calling/eb_cellid/pseudobulk_tmm/basic/8pcs/mash_fitted_model.full.rds")
```

```{r}
mash_colors <- color_map %>%
  arrange(type) %>%
  pull(color)
```

```{r}
mash_plot_meta(full_m, "SH3PXD2B_rs10042482", labels=colnames(full_m$result$lfsr),
               colors=meta.colors(box=mash_colors))
```

How does this compare to the effects we learned in the celltype-by-celltype analysis?
```{r fig.width=4, fig.height=5}
nonmash_viz <- filter(nonmash_effects, (variant_id=="rs10042482") & (phenotype_id=="SH3PXD2B"))

nonmash_colors <- color_map %>%
  right_join(nonmash_viz, by=c("type"="celltype")) %>%
  arrange(type) %>%
  pull(color)

color_map_mash_fig <- color_map %>%
  mutate(layer=factor(layer, levels=c("epithelium", "immune", "ecto_1", "ecto_2", "endoderm", "Squa_Epi", "mesoderm", "Vasc_Endo"))) %>%
  arrange(layer)

type_order <- color_map_mash_fig$type
color_order <- color_map_mash_fig$color
```


```{r}
mash_posterior_betas <- get_pm(full_m)["SH3PXD2B_rs10042482",]
mash_posterior_sds <- get_psd(full_m)["SH3PXD2B_rs10042482",]
mash_precision <- 1 / mash_posterior_sds ** 2
mash_celltypes <- colnames(get_pm(full_m))
mash_ests <- tibble(celltype=mash_celltypes,
                    beta=mash_posterior_betas,
                    sd=mash_posterior_sds,
                    precision=mash_precision,
                    approach=rep("mash", length(mash_celltypes)))

nonmash_betas <- nonmash_viz$slope
nonmash_sds <- nonmash_viz$slope_se
nonmash_precision <- 1 / nonmash_sds ** 2
nonmash_celltypes <- nonmash_viz$celltype
nonmash_ests <- tibble(celltype=nonmash_celltypes,
                    beta=nonmash_betas,
                    sd=nonmash_sds,
                    precision=nonmash_precision,
                    approach=rep("non-mash", length(nonmash_celltypes)))

mash_comp_viz <- bind_rows(mash_ests, nonmash_ests) %>%
  mutate(approach=case_match(factor(approach, levels=c("non-mash", "mash")), "non-mash" ~ "Univariate Analysis", "mash" ~ "Multivariate Analysis")) %>%
  mutate(approach=factor(approach, levels=c("Univariate Analysis", "Multivariate Analysis"))) %>%
  mutate(celltype=factor(celltype, levels=rev(type_order)))

mash_example_plot <- ggplot(mash_comp_viz, aes(x=celltype, y=beta, color=celltype)) +
  geom_hline(aes(yintercept=0), linetype="dashed", color="red") +
  geom_errorbar(aes(ymin=beta - sd, ymax=beta + sd), color="#D5D5D5", width=0.5) +
  geom_point(aes(size=precision), shape=15) +
  theme_classic(base_size=20) +
  theme(legend.position = "none", axis.text.x = element_text(angle=45, hjust=1)) +
  scale_color_manual(values=rev(color_order)) +
  coord_flip() +
  facet_grid(cols=vars(approach)) +
  xlab("Cell Type") +
  ylab("Estimated Effect Size") 
```


```{r fig.height=8, fig.width=20}
layout <- "
  AAAABBB
  AAAABBB
"
mash_factors_all + mash_example_plot +
  plot_layout(design=layout)
```

