```{r}
library(tidyverse)
library(vroom)
library(qvalue)
```

## DYNAMIC EQTL OVERLAP
```{r}
eb_cm_interactions <- read_tsv("/project2/gilad/jpopp/ebQTL/results/dynamic_qtl_calling/eb_cmstages/pseudobulk_tmm/10clpcs/tensorqtl_interactions.all.tsv")

elorbany_cm_interactions <- read_tsv("/project2/gilad/jpopp/ebQTL/results/dynamic_qtl_calling/elorbany_cm16bins/pseudobulk_tmm/5clpcs/tensorqtl_interactions.all.tsv")
```
Apply bonferroni correction and get the significant hits
```{r}
bonferroni_adjust <- function(p, n) {
  min(1, n*p)
}

eb_cm_tophits <- eb_cm_interactions %>%
  add_count(phenotype_id) %>%
  group_by(phenotype_id) %>%
  slice_min(order_by=pval_gi, n=1, with_ties=F, na_rm=T) %>%
  mutate(pval_gi_bonf=map2_dbl(pval_gi, n, bonferroni_adjust))
eb_cm_tophits$bh_gi_bonf <- p.adjust(eb_cm_tophits$pval_gi_bonf, "BH")
```



## EXPRESSION OVERLAP
Get the median expression in each cell type from elorbany - made in #temp/get_tmm_alltypes.R
```{r}
elorbany <- vroom("/project2/gilad/jpopp/ebQTL/temp/elorbany_cmstages.pseudobulk_tmm.tsv")
eb <- vroom("/project2/gilad/jpopp/ebQTL/temp/eb_cmstages.pseudobulk_tmm.tsv")
```
Get the median expression in each cell type
```{r}
elorbany_median <- elorbany %>%
  pivot_longer(!gene, names_to="sample", values_to="expression") %>%
  separate(sample, into=c("donor", "stage")) %>%
  group_by(stage, gene) %>%
  summarize(median_expression=median(expression)) %>%
  pivot_wider(id_cols=gene, names_from=stage, values_from=median_expression)

eb_median <- eb %>%
  pivot_longer(!gene, names_to="sample", values_to="expression") %>%
  separate(sample, into=c("donor", "stage")) %>%
  group_by(stage, gene) %>%
  summarize(median_expression=median(expression)) %>%
  pivot_wider(id_cols=gene, names_from=stage, values_from=median_expression)
```

Merge the columns for easier comparison
```{r}
eb_median_comb <- eb_median %>%
  rename(EB_CM=CM, EB_IPSC=IPSC, EB_MESENDO=MESENDO, EB_MESO=MESO, EB_PROG=PROG)

elorbany_median_comb <- elorbany_median %>%
  rename(EL_CM=CM, EL_IPSC=IPSC, EL_MES=MES, EL_CMES=CMES, EL_PROG=PROG)

medians_all <- inner_join(eb_median_comb, elorbany_median_comb, by="gene")
```

```{r}
expression_pairwise <- cor(medians_all[,c("EB_IPSC", "EB_MESENDO", "EB_MESO", "EB_PROG", "EB_CM")],
                           medians_all[,c("EL_IPSC", "EL_MES", "EL_CMES", "EL_PROG", "EL_CM")],
                           method="spearman", use="pairwise.complete.obs")
corrplot(expression_pairwise,
         method="color", addCoef.col = "black",
         is.corr=F)
```

## CIS EQTL OVERLAP
```{r}
eb_mash_fit <- readRDS("results/static_qtl_calling/eb_cmstages/pseudobulk_tmm/basic/5pcs/mash_fitted_model.tophits.rds")
elorbany_mash_fit <- readRDS("results/static_qtl_calling/elorbany_cmstages/pseudobulk_tmm/basic/5pcs/mash_fitted_model.tophits.rds")
```
These seem insanely inflated, will have to figure that out later

```{r}
eb_tensorqtl <- read_tsv("results/static_qtl_calling/eb_cmstages/pseudobulk_tmm/basic/8pcs/tensorqtl_permutations.all.tsv")

elorbany_tensorqtl <- read_tsv("results/static_qtl_calling/elorbany_cmstages/pseudobulk_tmm/basic/5pcs/tensorqtl_permutations.all.tsv")
```

Nothing was significant at a single-stage level in Elorbany, so we're going to focus on just our EB hits
```{r}
eb_qtl_effect_sizes <- eb_tensorqtl %>%
  filter(q <= 0.05) %>%
  mutate(context=paste0("EB_", context)) %>%
  unite(gv, phenotype_id, variant_id, sep="_") %>%
  pivot_wider(id_cols=gv, names_from=context, values_from=slope)
```

Now let's see what the Elorbany data looks like there
```{r}
elorbany_all <- read_tsv("results/static_qtl_calling/elorbany_cmstages/pseudobulk_tmm/basic/5pcs/tensorqtl_nominal.betas.tsv")
```

```{r}
effect_size_comp <- elorbany_all %>%
  rename_with(function(s){paste0("EL_", s)}, .cols=!gv) %>%
  right_join(eb_qtl_effect_sizes, by="gv")
```

```{r}
eqtl_pairwise <- cor(effect_size_comp[,c("EB_IPSC", "EB_MESENDO", "EB_MESO", "EB_PROG", "EB_CM")],
                     effect_size_comp[,c("EL_IPSC", "EL_MES", "EL_CMES", "EL_PROG", "EL_CM")],
                     method="spearman", use="pairwise.complete.obs")
corrplot(eqtl_pairwise,
         method="color", addCoef.col = "black",
         is.corr=F)
```
How many tests were evaluated to come up with these figures?
```{r}
both_non_na <- function(eb_type, elorbany_type, comp_df=effect_size_comp) {
  sum(!(is.na(comp_df[,eb_type]) | is.na(comp_df[,elorbany_type])))
}
eb_types <- c("EB_IPSC", "EB_MESENDO", "EB_MESO", "EB_PROG", "EB_CM")
el_types <- c("EL_IPSC", "EL_MES", "EL_CMES", "EL_PROG", "EL_CM")

num_tests_evaluated <- matrix(nrow=5, ncol=5, dimnames=list(eb_types, el_types))

for (eb_i in eb_types) {
  for (el_i in el_types) {
    num_tests_evaluated[eb_i, el_i] <- both_non_na(eb_i, el_i)
  }
}

corrplot(num_tests_evaluated,
         method="color", addCoef.col = "black",
         is.corr=F)
```



```{r}

elorbany_ipscloci <- elorbany_all %>%
  filter(gv %in% ipsc_loci)

ipsc_comparison <- ipsc_effects %>%
  select(gv, slope) %>%
  right_join(elorbany_ipscloci, by=join_by(gv==gv))
```

How about among the very few PROG hits
```{r}
late_effects <- eb_qtl_hits %>%
  filter(context %in% c("PROG", "CM")) %>%
  select(phenotype_id, variant_id, slope) %>%
  mutate(gv=paste0(phenotype_id, "_", variant_id))
```

```{r}
elorbany_lateloci <- elorbany_all %>%
  filter(gv %in% late_effects$gv)

late_comparison <- late_effects %>%
  select(gv, slope) %>%
  right_join(elorbany_lateloci, by=join_by(gv==gv))

```

```{r}
cor(late_comparison$slope[!is.na(late_comparison$IPSC)], late_comparison$IPSC[!is.na(late_comparison$IPSC)])
cor(late_comparison$slope[!is.na(late_comparison$MES)], late_comparison$MES[!is.na(late_comparison$MES)])
cor(late_comparison$slope[!is.na(late_comparison$CMES)], late_comparison$CMES[!is.na(late_comparison$CMES)])
cor(late_comparison$slope[!is.na(late_comparison$PROG)], late_comparison$PROG[!is.na(late_comparison$PROG)])
cor(late_comparison$slope[!is.na(late_comparison$CM)], late_comparison$CM[!is.na(late_comparison$CM)])
```

How about among the middle
```{r}
mid_effects <- eb_qtl_hits %>%
  filter(context %in% c("MESENDO", "MESO")) %>%
  select(phenotype_id, variant_id, slope) %>%
  mutate(gv=paste0(phenotype_id, "_", variant_id))
```

```{r}
elorbany_midloci <- elorbany_all %>%
  filter(gv %in% mid_effects$gv)

mid_comparison <- mid_effects %>%
  select(gv, slope) %>%
  right_join(elorbany_midloci, by=join_by(gv==gv))

```

```{r}
cor(mid_comparison$slope[!is.na(mid_comparison$IPSC)], mid_comparison$IPSC[!is.na(mid_comparison$IPSC)])
cor(mid_comparison$slope[!is.na(mid_comparison$MES)], mid_comparison$MES[!is.na(mid_comparison$MES)])
cor(mid_comparison$slope[!is.na(mid_comparison$CMES)], mid_comparison$CMES[!is.na(mid_comparison$CMES)])
cor(mid_comparison$slope[!is.na(mid_comparison$PROG)], mid_comparison$PROG[!is.na(mid_comparison$PROG)])
cor(mid_comparison$slope[!is.na(mid_comparison$CM)], mid_comparison$CM[!is.na(mid_comparison$CM)])
```
