```{r}
library(tidyverse)
library(vroom)
library(qvalue)
library(MatrixGenerics)
library(RColorBrewer)
```

## Cardiomyocyte Trajectory
```{r}
cm_early <- vroom("results/static_eqtl_followup/qtl_sets/dynamic-eqtls/earlydynamic-eb-cm-signif_variant_gene_pairs.bed")

cm_early_overlap <- vroom("results/static_eqtl_followup/qtl_sets/dynamic-eqtls/earlydynamic-eb-cm-signif_variant_gene_pairs.all_tissue_overlap.bed",
                   col_names=c("#CHR", "START", "END", "EB_ENSG", "EB_HGNC", "EB_VARIANT_ID", "EB_CLASS_TRAJ",
                               "GTEX_CHR", "GTEX_START", "GTEX_END", "GTEX_ENSG", "REF", "ALT"))

cm_late <- vroom("results/static_eqtl_followup/qtl_sets/dynamic-eqtls/latedynamic-eb-cm-signif_variant_gene_pairs.bed")

cm_late_overlap <- vroom("results/static_eqtl_followup/qtl_sets/dynamic-eqtls/latedynamic-eb-cm-signif_variant_gene_pairs.all_tissue_overlap.bed",
                   col_names=c("#CHR", "START", "END", "EB_ENSG", "EB_HGNC", "EB_VARIANT_ID", "EB_CLASS_TRAJ",
                               "GTEX_CHR", "GTEX_START", "GTEX_END", "GTEX_ENSG", "REF", "ALT"))

cm_switch <- vroom("results/static_eqtl_followup/qtl_sets/dynamic-eqtls/switchdynamic-eb-cm-signif_variant_gene_pairs.bed")
cm_switch_overlap <- vroom("results/static_eqtl_followup/qtl_sets/dynamic-eqtls/switchdynamic-eb-cm-signif_variant_gene_pairs.all_tissue_overlap.bed",
                   col_names=c("#CHR", "START", "END", "EB_ENSG", "EB_HGNC", "EB_VARIANT_ID", "EB_CLASS_TRAJ",
                               "GTEX_CHR", "GTEX_START", "GTEX_END", "GTEX_ENSG", "REF", "ALT"))
```

How many early dynamic eQTLs overlap with GTEx?
```{r}
cm_early_dynamic_egenes <- length(unique(cm_early$EB_HGNC))
cm_early_dynamic_egenes_unique <- length(setdiff(unique(cm_early$EB_HGNC), unique(cm_early_overlap$EB_HGNC)))
```

How many late dynamic eQTLs overlap with GTEx?
```{r}
cm_late_dynamic_egenes <- length(unique(cm_late$EB_HGNC))
cm_late_dynamic_egenes_unique <- length(setdiff(unique(cm_late$EB_HGNC), unique(cm_late_overlap$EB_HGNC)))
```

## Neuronal Trajectory
```{r}
neur_early <- vroom("results/static_eqtl_followup/qtl_sets/dynamic-eqtls/earlydynamic-eb-neur-signif_variant_gene_pairs.bed")

neur_early_overlap <- vroom("results/static_eqtl_followup/qtl_sets/dynamic-eqtls/earlydynamic-eb-neur-signif_variant_gene_pairs.all_tissue_overlap.bed",
                   col_names=c("#CHR", "START", "END", "EB_ENSG", "EB_HGNC", "EB_VARIANT_ID", "EB_CLASS_TRAJ",
                               "GTEX_CHR", "GTEX_START", "GTEX_END", "GTEX_ENSG", "REF", "ALT"))

neur_late <- vroom("results/static_eqtl_followup/qtl_sets/dynamic-eqtls/latedynamic-eb-neur-signif_variant_gene_pairs.bed")

neur_late_overlap <- vroom("results/static_eqtl_followup/qtl_sets/dynamic-eqtls/latedynamic-eb-neur-signif_variant_gene_pairs.all_tissue_overlap.bed",
                   col_names=c("#CHR", "START", "END", "EB_ENSG", "EB_HGNC", "EB_VARIANT_ID", "EB_CLASS_TRAJ",
                               "GTEX_CHR", "GTEX_START", "GTEX_END", "GTEX_ENSG", "REF", "ALT"))

neur_switch <- vroom("results/static_eqtl_followup/qtl_sets/dynamic-eqtls/switchdynamic-eb-neur-signif_variant_gene_pairs.bed")
neur_switch_overlap <- vroom("results/static_eqtl_followup/qtl_sets/dynamic-eqtls/switchdynamic-eb-neur-signif_variant_gene_pairs.all_tissue_overlap.bed",
                   col_names=c("#CHR", "START", "END", "EB_ENSG", "EB_HGNC", "EB_VARIANT_ID", "EB_CLASS_TRAJ",
                               "GTEX_CHR", "GTEX_START", "GTEX_END", "GTEX_ENSG", "REF", "ALT"))
```

How many early dynamic eQTLs overlap with GTEx?
```{r}
neur_early_dynamic_egenes <- length(unique(neur_early$EB_HGNC))
neur_early_dynamic_egenes_unique <- length(setdiff(unique(neur_early$EB_HGNC), unique(neur_early_overlap$EB_HGNC)))
```

How many late dynamic eQTLs overlap with GTEx?
```{r}
neur_late_dynamic_egenes <- length(unique(neur_late$EB_HGNC))
neur_late_dynamic_egenes_unique <- length(setdiff(unique(neur_late$EB_HGNC), unique(neur_late_overlap$EB_HGNC)))
```

## Hepatoblast Trajectory
```{r}
hep_early <- vroom("results/static_eqtl_followup/qtl_sets/dynamic-eqtls/earlydynamic-eb-hep-signif_variant_gene_pairs.bed")

hep_early_overlap <- vroom("results/static_eqtl_followup/qtl_sets/dynamic-eqtls/earlydynamic-eb-hep-signif_variant_gene_pairs.all_tissue_overlap.bed",
                   col_names=c("#CHR", "START", "END", "EB_ENSG", "EB_HGNC", "EB_VARIANT_ID", "EB_CLASS_TRAJ",
                               "GTEX_CHR", "GTEX_START", "GTEX_END", "GTEX_ENSG", "REF", "ALT"))

hep_late <- vroom("results/static_eqtl_followup/qtl_sets/dynamic-eqtls/latedynamic-eb-hep-signif_variant_gene_pairs.bed")

hep_late_overlap <- vroom("results/static_eqtl_followup/qtl_sets/dynamic-eqtls/latedynamic-eb-hep-signif_variant_gene_pairs.all_tissue_overlap.bed",
                   col_names=c("#CHR", "START", "END", "EB_ENSG", "EB_HGNC", "EB_VARIANT_ID", "EB_CLASS_TRAJ",
                               "GTEX_CHR", "GTEX_START", "GTEX_END", "GTEX_ENSG", "REF", "ALT"))

hep_switch <- vroom("results/static_eqtl_followup/qtl_sets/dynamic-eqtls/switchdynamic-eb-hep-signif_variant_gene_pairs.bed")
hep_switch_overlap <- vroom("results/static_eqtl_followup/qtl_sets/dynamic-eqtls/switchdynamic-eb-hep-signif_variant_gene_pairs.all_tissue_overlap.bed",
                   col_names=c("#CHR", "START", "END", "EB_ENSG", "EB_HGNC", "EB_VARIANT_ID", "EB_CLASS_TRAJ",
                               "GTEX_CHR", "GTEX_START", "GTEX_END", "GTEX_ENSG", "REF", "ALT"))

```

How many early dynamic eQTLs overlap with GTEx?
```{r}
hep_early_dynamic_egenes <- length(unique(hep_early$EB_HGNC))
hep_early_dynamic_egenes_unique <- length(setdiff(unique(hep_early$EB_HGNC), unique(hep_early_overlap$EB_HGNC)))
```

How many late dynamic eQTLs overlap with GTEx?
```{r}
hep_late_dynamic_egenes <- length(unique(hep_late$EB_HGNC))
hep_late_dynamic_egenes_unique <- length(setdiff(unique(hep_late$EB_HGNC), unique(hep_late_overlap$EB_HGNC)))
```

## Combined summary
```{r}
all_dynamic_eqtls <- bind_rows(
  select(cm_early, c(EB_HGNC, EB_VARIANT_ID, EB_CLASS_TRAJ)),
  select(neur_early, c(EB_HGNC, EB_VARIANT_ID, EB_CLASS_TRAJ)),
  select(hep_early, c(EB_HGNC, EB_VARIANT_ID, EB_CLASS_TRAJ)),
  select(cm_switch, c(EB_HGNC, EB_VARIANT_ID, EB_CLASS_TRAJ)),
  select(neur_switch, c(EB_HGNC, EB_VARIANT_ID, EB_CLASS_TRAJ)),
  select(hep_switch, c(EB_HGNC, EB_VARIANT_ID, EB_CLASS_TRAJ)),
  select(cm_late, c(EB_HGNC, EB_VARIANT_ID, EB_CLASS_TRAJ)),
  select(neur_late, c(EB_HGNC, EB_VARIANT_ID, EB_CLASS_TRAJ)),
  select(hep_late, c(EB_HGNC, EB_VARIANT_ID, EB_CLASS_TRAJ))
) %>%
  separate(EB_CLASS_TRAJ, into=c("eb", "traj", "class")) %>%
  select(-c(eb))
```

Identify any classification conflicts
```{r}
classification_conflicts <- all_dynamic_eqtls %>%
  distinct(EB_HGNC, EB_VARIANT_ID, class) %>%
  add_count(EB_HGNC, EB_VARIANT_ID) %>%
  filter(n > 1) %>%
  arrange(EB_HGNC, EB_VARIANT_ID, class, n)
```
PIGP has one early and one late eQTL at rs2298682, TMED2 and FAAP20 have late/ switch effects at rs112931584 and rs262693 respectively (randomly selected from the blocks)

```{r}
all_dynamic_eqtls_noconflict <- anti_join(all_dynamic_eqtls, classification_conflicts, by=c("EB_HGNC"))
distinct_dynamic_eqtls_noconflict <- distinct(all_dynamic_eqtls_noconflict, EB_HGNC, EB_VARIANT_ID, class)

dplyr::count(distinct_dynamic_eqtls_noconflict, class)
```

What fraction of dynamic eQTLs display GTEx overlap?
```{r}
all_gtex_overlaps_noconflict <- bind_rows(
  select(cm_early_overlap, c(EB_HGNC, EB_VARIANT_ID, EB_CLASS_TRAJ)),
  select(neur_early_overlap, c(EB_HGNC, EB_VARIANT_ID, EB_CLASS_TRAJ)),
  select(hep_early_overlap, c(EB_HGNC, EB_VARIANT_ID, EB_CLASS_TRAJ)),
  select(cm_switch_overlap, c(EB_HGNC, EB_VARIANT_ID, EB_CLASS_TRAJ)),
  select(neur_switch_overlap, c(EB_HGNC, EB_VARIANT_ID, EB_CLASS_TRAJ)),
  select(hep_switch_overlap, c(EB_HGNC, EB_VARIANT_ID, EB_CLASS_TRAJ)),
  select(cm_late_overlap, c(EB_HGNC, EB_VARIANT_ID, EB_CLASS_TRAJ)),
  select(neur_late_overlap, c(EB_HGNC, EB_VARIANT_ID, EB_CLASS_TRAJ)),
  select(hep_late_overlap, c(EB_HGNC, EB_VARIANT_ID, EB_CLASS_TRAJ))
) %>%
  separate(EB_CLASS_TRAJ, into=c("eb", "traj", "class")) %>%
  select(-c(eb)) %>%
  anti_join(classification_conflicts, by=c("EB_HGNC"))

distinct_gtex_overlaps_noconflict <- distinct(all_gtex_overlaps_noconflict, EB_HGNC, EB_VARIANT_ID, class)

nrow(distinct_gtex_overlaps_noconflict) / nrow(all_dynamic_eqtls_noconflict)
```

Break this down by classification
```{r}
early_dynamic_eqtls_noconflict <- filter(distinct_dynamic_eqtls_noconflict, class=="early")
late_switch_dynamic_eqtls_noconflict <- filter(distinct_dynamic_eqtls_noconflict, class %in% c("late", "switch"))

early_gtex_overlaps_noconflict <- filter(distinct_gtex_overlaps_noconflict, class=="early")
late_switch_gtex_overlaps_noconflict <- filter(distinct_gtex_overlaps_noconflict, class %in% c("late", "switch"))

nrow(early_gtex_overlaps_noconflict) / nrow(early_dynamic_eqtls_noconflict)
nrow(late_switch_gtex_overlaps_noconflict) / nrow(late_switch_dynamic_eqtls_noconflict)
```


What fraction of genes with an early eQTL display GTEx overlap?
```{r}

early_dynamic_eqtls_noconflict_nogtex <- early_dynamic_eqtls_noconflict %>%
  anti_join(cm_early_overlap, by=c("EB_HGNC")) %>%
  anti_join(neur_early_overlap, by=c("EB_HGNC")) %>%
  anti_join(hep_early_overlap, by=c("EB_HGNC"))

length(unique(early_dynamic_eqtls_noconflict_nogtex$EB_HGNC)) / length(unique(early_dynamic_eqtls_noconflict$EB_HGNC))
```

```{r}
late_dynamic_eqtls_noconflict <- filter(distinct_dynamic_eqtls_noconflict, class=="late")
late_dynamic_eqtls_noconflict_nogtex <- late_dynamic_eqtls_noconflict %>%
  anti_join(cm_late_overlap, by=c("EB_HGNC")) %>%
  anti_join(neur_late_overlap, by=c("EB_HGNC")) %>%
  anti_join(hep_late_overlap, by=c("EB_HGNC"))

length(unique(late_dynamic_eqtls_noconflict_nogtex$EB_HGNC)) / length(unique(late_dynamic_eqtls_noconflict$EB_HGNC))
```